---
title: "Minimum metadata recommendations"
format: html
---
### Background 

Metadata describing every dataset, data products, and written article is needed to ensure that these resources can be found, accessed, and reused. If you are publishing data in a robust repository, these metadata recommendations may be met or exceeded by those of the repository.  If not, these recommendations can give a guide as to what types of information should accompany and dataset or research output.  These essential metadata are meant to apply across all research domains and subject areas, and for all research outputs.  

### References used to develop these guidelines

[MEDIN Discovery Metadata Standard](https://medin.org.uk/sites/medin/files/documents/MEDIN_Schema_Documentation_3_1_2_brief.pdf)  
[Creative Commons TASL](https://wiki.creativecommons.org/wiki/Recommended_practices_for_attribution)  

### Other metadata resources

[DCAT-US](https://resources.data.gov/resources/dcat-us/)   
[EML](https://eml.ecoinformatics.org/)  
[ISO19115-2](https://www.iso.org/obp/ui/#iso:std:iso:19115:-2:ed-2:v1:en)  


```{r, echo=FALSE, warning=FALSE, include=FALSE}
library(tidyverse); library(knitr); library(kableExtra)

table1 <- read.csv("metadata_table.csv", header = TRUE, stringsAsFactors = FALSE)
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
table1[6,2] <- paste0("Terms that describe how the research output (including data) can be shared and reused.  Good practice is to explicitly state the license selected for the data.  The recommended licenses are ","<a href=\"","https://creativecommons.org/publicdomain/zero/1.0/","\">","CC0","</a>", " (public domain) or ", "<a href=\"","https://creativecommons.org/licenses/by/4.0/","\">","CC-BY","</a>", ".")

table1[5,2] <- paste0("A unique persistent identifier such as a ", "<a href=\"","https://www.doi.org/the-identifier/what-is-a-doi/","\">","Digital Object Identifier","</a>", " (DOI)")

table1[12,2] <- paste0("Name and persistent ID (e.g. ", "<a href=\"","https://ror.org/","\">","ROR","</a>", ", if available) of the organization(s) funding this research")
                    
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#| html-table-processing: none

TBL <- kbl(table1, booktabs = TRUE, align = ("llccc"), format = "html", escape = FALSE,
           col.names = c("Minimum Metadata", 
                         "Common-language description of metadata terms", 
                         "EML", "ISO19115-2", "DCAT-US")) %>% 
       kable_styling(full_width = FALSE, position = "center", fixed_thead = TRUE) %>% 
       add_header_above(header = c(" " = 1, "Description" = 1, "Equivalent fields in common metadata standards" = 3), 
                        align = "l") %>% 
       column_spec(1, width = "12em") %>% 
       column_spec(2, width = "45em") %>% 
       column_spec(3:5, width = "8em") 
      
TBL
```



